---
title: "run_scType"
Author: 'Joan Kant'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{run_scType}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

## Download the `pbmc3k` dataset

For this tutorial, you'll use a PBMC 3k dataset (see Seurat tutorial for more details on how to load the dataset using Seurat, [pbmc3k tutorial](https://satijalab.org/seurat/articles/pbmc3k_tutorial.html))

There are two options to obtain the dataset: 

1. Download the gene-count matrix [here](https://cf.10xgenomics.com/samples/cell/pbmc3k/pbmc3k_filtered_gene_bc_matrices.tar.gz) and create a Seurat object
2. Use the SeuratData R-package 

### Option 1 

```{r}
# load libraries
library(Seurat)

# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = "./filtered_gene_bc_matrices/hg19/")
# Initialize the Seurat object with the raw (non-normalized data).
pbmc <- CreateSeuratObject(counts = pbmc.data, project = "pbmc3k", min.cells = 3, min.features = 200)
```

### Option 2: Use the SeuratData R-package

```{r install-pbmc3k eval = FALSE}
# 1. install SeuratData
pak::pkg_install("satijalab/seurat-data")

# 2. Install the 'pbmc3k' dataset
library(SeuratData)
InstallData("pbmc3k")

# 3. Load the dataset -> seurat object
data(pbmc3k)
```

## Prepare dataset

Normalize and cluster the data. 

> `pbmc` is the Seurat object

```{r}
library(Seurat)
pbmc <- pbmc3k
pbmc <- UpdateSeuratObject(pbmc)
# normalize data
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")
# pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5) # make some filtering based on QC metrics visualizations, see Seurat tutorial: https://satijalab.org/seurat/articles/pbmc3k_tutorial.html
pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)

# scale and run PCA
pbmc <- ScaleData(pbmc, features = rownames(pbmc))
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))

# Check number of PC components (we selected 10 PCs for downstream analysis, based on Elbow plot)
print(ElbowPlot(pbmc))

# cluster and visualize
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.8)
pbmc <- RunUMAP(pbmc, dims = 1:10)
print(DimPlot(pbmc, reduction = "umap"))
```

## Run scType


```{r setup}
library(sctype)
data(ScTypeDB_full)

pbmc <- run_sctype(pbmc,
    assay = "RNA",
    scaled = TRUE,
    known_tissue_type = "Immune system",
    custom_marker_file = ScTypeDB_full,
    name = "sctype_classification",
    plot_umap = TRUE, plot_bubble = TRUE
)
```